### 2019-01-31 texs/paper_base.tex
toriaezu dasita

\date{December 11, 2018}
%-------------------

@@

@@

% (TODO)LSTM,Seq2seqの説明、その後Attentionについて説明、さらにその後seq2treeについて説明する。
\subsection{RNN and LSTM}
A recurrent neural network (RNN) is a neural network for dealing with sequence data.
The structure of the RNN is the sequence of the same node network units, as shown in Figre~\ref{fig:encoder} and Figre~\ref{fig:decoder}. 
The long short-term memory (LSTM) is a unit of the RNN which we used in the experiment. 
It was first introduced by \sloppy \citet{first_LSTM} , for overcoming the vanishing gradient problem and the exploding gradient problem.
In short, an LSTM unit receives two input vectors, 
one is the internal state of the LSTM unit, and the other is the input for the LSTM unit, 
and the LSTM unit emits one output vector, which represents the next internal state of the LSTM unit.

@@

\begin{figure}[]
	\begin{center}
	\includegraphics[height=4cm]{encoder.png}
	\end{center}
	\caption{The encoder network of the sequence-to-sequence model}
	\label{fig:encoder}
	\begin{center}
	\includegraphics[height=5.7cm]{decoder.png}
	\end{center}
	\caption{ The decoder network of the sequence-to-sequence model}
	\label{fig:decoder}
\end{figure}

@@

As the mentioned in the previous section, the pure sequence-to-sequence model uses the output of the decoder network $ \mathrm{dec}_{k} $ for estimating 
the distribution $ p(y_k|x,[y_1,\dots,y_{k-1}]) $.
In the sequence-to-sequence model with attention, 
a context vector $ \mathrm{ctx}_{k} $ is calculalted from $ \mathrm{dec}_{k} $ and the internal states of encoder networks $ \mathrm{enc}_{1} \dots \mathrm{enc}_{n} $, 
then the distribution is estimated from the output vector $ \mathrm{dec}_{k} $ and context vector $ \mathrm{ctx}_{k} $.
The context vector $ \mathrm{ctx}_{k} $ is calculated as follows.
For each encoder state $ \mathrm{enc}_{l} $, the importance of the state is calculated as $ a_{l} = \mathrm{score}(\mathrm{dec}_{k},\mathrm{enc}_{l}) $,
then $ \mathrm{ctx}_{k} $ is calculated as $ \mathrm{ctx}_{k} = \sum_{l=1}^{n} \frac{\exp{a_{l}}}{\sum_{l=1}^{n}\exp{a_{l}}} \mathrm{enc}_{l} $, 
as the weighted average of $ \mathrm{enc}_{1} \dots \mathrm{enc}_{n} $.
% Multiple score functions are introduced by \citet{dot_attention}, and we choiced 
In \citet{dot_attention}, various kinds of score functions were suggested.
In our experiment, we used a general score function, which is calculated as $ \mathrm{score}(\mathrm{dec}_{k},\mathrm{enc}_{l}) = \mathrm{dec}_{k}^T W \mathrm{enc}_{l} $ ,
where $ W $ is a parameter matrix.

@@

% In the pure sequence-to-sequence model, the conditional probability  $ p(y_k|x,[y_1,\dots,y_{k-1}]) $ is calculated from $ dec_{k} $ ,which is the internal state of the decoder network.
% In the sequence-to-sequence with attention model, the probability is calcutated from $ dec_{k} $ and the context vevctor $ c_{k} $.
% $ c_{k} $

@@

The decoder network is a network for generating tree structured output. 
A tree structure is representable by a sequence of the actions which expands a nonterminal symbol of the constructing tree.
For example, the tree shown in Figure~\ref{fig:ast_zu} is generated from the sequence $ [\mathrm{act}_1, \dots, \mathrm{act}_{10}] $, 
where action $ \mathrm{act}_k $ expands the left-most nonterminal symbol in the tree generated from the action sequence $ [\mathrm{act}_1, \dots, \mathrm{act}_{k-1}] $. 
Suppose the tree $t$ is generated by the action sequence $ [\mathrm{act}_1, \dots, \mathrm{act}_m] $, 
then the tree-generating probability $ p(t|x) $ is decomposed into $ \prod_{k=1}^n p(\mathrm{act}_k|x,[\mathrm{act}_1, \dots, \mathrm{act}_{k-1}]) $. 
Thus the sequence-to-sequence model is now applicable to generating tree structure, by embeding each action $\mathrm{act}_k$ to represent vector $\mathrm{av}_k$ and 
calculate $ \mathrm{lstm}_{\mathrm{dec}}(\mathrm{dec}_{k},\mathrm{av}_{k}) $ sequentially for estimating $p(\mathrm{act}_{k+1}|x,[\mathrm{act}_1, \dots, \mathrm{act}_{k}]) $.

@@

These pairs were generated as follows. Let S0 be an original C source code.
\begin{enumerate}
\item Remove all preprocessors, such as \texttt{\#include}, in S0 to generate preprocessed C code S1. 
\item Tokenize S1 and rearrange into S2 so that there is exactly one token for each line. 
\item 
Compile S2 by the GNU C compiler with no optimization option (-O0) and debug option (-g) to get assembly code S3. 
With the debug option, S3 contains the information about the corresponding source code positions in S2. 
% which each assembly operations are originated from.
\item Assemble S3 to object file S4 and disassemble S4 to get information eliminated assembly code S5.
\item 
Parse S2 with a Clang compiler to get the abstract syntax tree (AST) of the source code and enumerate the subtrees of the AST.
Each subtree of the AST has the correspoinding source code positions in S2.
Therefore, we can get a pair of an assembly code and a C fragment corresponding to the subtree 
by using the informations in S2, S3 and S5.
\end{enumerate}

@@

\subsection{Assembly Code Preprocess}
% 155 words
There are two ways for preprocessing assembly codes.

@@

The second way is converting each code into the tokenized sequence of assembly codes. 
That is, we split the disassembled instructions in S5 into the sequence of symbols, 
concatenate these sequences to one sequence, 
and modify it as follows:
\begin{itemize}
\item For each end of the instructions, the \texttt{ENDLINE} token is added to indicate the end of the instructions.
\item 
The relative addressing, such as the address of the \texttt{call} or \texttt{jump} instructions, 
is relabeled similarly to $\alpha$-conversion for making the code position independent and for reducing the vocabulary size.
\item Big integer values are clipped. 
More specifically, an integer in the range from -255 to 256 are preserved as it is, and other integers are clipped and substituted with the \texttt{\_\_HEX\_\_} symbol.
This also reduces the vocabulary size.
\end{itemize}

@@

\subsection{C Code Preprocess}
% 232 words
The C source codes were parsed to AST by a Clang compiler \citep{clang} and its Python bindings together with pycparser \citep{pycparser}, 
because all of the tools are individually not sufficient for generating a dataset.
Then, the AST is converted to input data as follows:

\begin{itemize}
\item The variable-ary nodes and optional nodes were converted. 
For example, the \texttt{CompoundStatement} node has a variable number of statements. 
We converted such a list of nodes to the chains of the node \texttt{CONS} and the node \texttt{NIL}, as the list representation in a functional language.
Another example is the \texttt{if} statement, whose \texttt{else} clause is sometimes missing. 
We converted such optional nodes to the alternative of the \texttt{SOME} node and the \texttt{NONE} node, as the optional data structure in a functional language.
\item The variable names, function names, label names, or type names were renamed like $\alpha$-conversion.
\item There are four type of literals in the C language. 
An integer literal less than 256 was preserved, and an integer literal greater than 256 was clipped to 256.
The string, float and character literals were converted to the \texttt{\_\_LITERAL\_\_} symbol.

@@

\section{Evaluation Metrics}
We evaluated the quality of the results by two metrics, 
the edit distance ratio and the bilingual evaluation understudy (BLEU) \citep{BLEU}.

The edit distance ratio is a metric for for the correctness of the decompilation result.
Here, the edit distance ratio between two strings means the Levenshtein distance \citep{levensthein_dist} divided by the average length of the two strings.
Ideally, the correctness of the decompilation result should be measured by the distance of the meaning of the program.
However the decompilation result is usually uncompilable, 
thus we used the edit distance of the source codes as an approximation of the correctness.

>>>>

\date{February 1, 2018}
%-------------------

@@

\tableofcontents

@@

% (TODO)LSTM,Seq2seqの説明、その後Attentionについて説明、さらにその後seq2treeについて説明する。
\subsection{RNN and LSTM}
A recurrent neural network (RNN) is a neural network for dealing with sequence data.
The structure of the RNN is the sequence of the same node network units, as shown in Figure~\ref{fig:encoder} and Figure~\ref{fig:decoder}. 
The long short-term memory (LSTM) is a unit of the RNN which we used in the experiment. 
It was first introduced by \sloppy \citet{first_LSTM} , for overcoming the vanishing gradient problem and the exploding gradient problem.
In short, an LSTM unit receives two input vectors, 
one is the internal state of the LSTM unit, and the other is the input for the LSTM unit, 
and the LSTM unit emits one output vector, which represents the next internal state of the LSTM unit.

@@

\begin{figure}[]
	\begin{center}
	\includegraphics[height=3cm]{encoder.png}
	\end{center}
	\caption{The encoder network of the sequence-to-sequence model}
	\label{fig:encoder}
	\begin{center}
	\includegraphics[height=4.2cm]{decoder.png}
	\end{center}
	\caption{ The decoder network of the sequence-to-sequence model}
	\label{fig:decoder}
\end{figure}

@@

As the mentioned in the previous section, the pure sequence-to-sequence model uses the output of the decoder network $ \mathrm{dec}_{k} $ for estimating 
the distribution $ p(y_k|x,[y_1,\dots,y_{k-1}]) $.
In the sequence-to-sequence model with attention, 
a context vector $ \mathrm{ctx}_{k} $ is calculated from $ \mathrm{dec}_{k} $ and the internal states of encoder networks $ \mathrm{enc}_{1} \dots \mathrm{enc}_{n} $, 
then the distribution is estimated from the output vector $ \mathrm{dec}_{k} $ and context vector $ \mathrm{ctx}_{k} $.
The context vector $ \mathrm{ctx}_{k} $ is calculated as follows.
For each encoder state $ \mathrm{enc}_{l} $, the importance of the state is calculated as $ a_{l} = \mathrm{score}(\mathrm{dec}_{k},\mathrm{enc}_{l}) $,
then $ \mathrm{ctx}_{k} $ is calculated as $ \mathrm{ctx}_{k} = \sum_{l=1}^{n} \frac{\exp{a_{l}}}{\sum_{l=1}^{n}\exp{a_{l}}} \mathrm{enc}_{l} $, 
as the weighted average of $ \mathrm{enc}_{1} \dots \mathrm{enc}_{n} $.
% Multiple score functions are introduced by \citet{dot_attention}, and we choiced 
In \citet{dot_attention}, various kinds of score functions were suggested.
In our experiment, we used a general score function, which is calculated as $ \mathrm{score}(\mathrm{dec}_{k},\mathrm{enc}_{l}) = \mathrm{dec}_{k}^T W \mathrm{enc}_{l} $ ,
where $ W $ is a parameter matrix.

@@

% In the pure sequence-to-sequence model, the conditional probability  $ p(y_k|x,[y_1,\dots,y_{k-1}]) $ is calculated from $ dec_{k} $ ,which is the internal state of the decoder network.
% In the sequence-to-sequence with attention model, the probability is calculated from $ dec_{k} $ and the context vevctor $ c_{k} $.
% $ c_{k} $

@@

The decoder network is a network for generating tree structured output. 
A tree structure is representable by a sequence of the actions which expands a nonterminal symbol of the constructing tree.
For example, the tree shown in Figure~\ref{fig:ast_zu} is generated from the sequence $ [\mathrm{act}_1, \dots, \mathrm{act}_{10}] $, 
where action $ \mathrm{act}_k $ expands the left-most nonterminal symbol in the tree generated from the action sequence $ [\mathrm{act}_1, \dots, \mathrm{act}_{k-1}] $. 
Suppose the tree $t$ is generated by the action sequence $ [\mathrm{act}_1, \dots, \mathrm{act}_m] $, 
then the tree-generating probability $ p(t|x) $ is decomposed into $ \prod_{k=1}^n p(\mathrm{act}_k|x,[\mathrm{act}_1, \dots, \mathrm{act}_{k-1}]) $. 
Thus the sequence-to-sequence model is now applicable to generating tree structure, by embedding each action $\mathrm{act}_k$ to represent vector $\mathrm{av}_k$ and 
calculate $ \mathrm{lstm}_{\mathrm{dec}}(\mathrm{dec}_{k},\mathrm{av}_{k}) $ sequentially for estimating $p(\mathrm{act}_{k+1}|x,[\mathrm{act}_1, \dots, \mathrm{act}_{k}]) $.

@@

These pairs were generated as follows. Let S0 be an original C source code.
\begin{enumerate}
\item Remove all preprocessors, such as \texttt{\#include}, in S0 to generate preprocessed C code S1. 
\item Tokenize S1 and rearrange into S2 so that there is exactly one token for each line. 
\item 
Compile S2 by the GNU C compiler with no optimization option (-O0) and debug option (-g) to get assembly code S3. 
With the debug option, S3 contains the information about the corresponding source code positions in S2. 
% which each assembly operations are originated from.
\item Assemble S3 to object file S4 and disassemble S4 to get information eliminated assembly code S5.
\item 
Parse S2 with a Clang compiler to get the abstract syntax tree (AST) of the source code and enumerate the subtrees of the AST.
Each subtree of the AST has the corresponding source code positions in S2.
Therefore, we can get a pair of an assembly code and a C fragment corresponding to the subtree 
by using the informations in S2, S3 and S5.
\end{enumerate}

@@

\subsection{Assembly Code Preprocessing}
% 155 words
There are two ways for preprocessing assembly codes.

@@

The second way is converting each code into the tokenized sequence of assembly codes. 
That is, we split the disassembled instructions in S5 into the sequence of symbols, 
concatenate these sequences to one sequence, 
and modify it as follows:
\begin{itemize}
\item For each end of the instructions, the \texttt{ENDLINE} token is added to indicate the end of the instructions.
\item 
The relative addressing, such as the address of the \texttt{call} or \texttt{jump} instructions, 
is normalized by renaming so as to make the code position independent and to reduce the vocabulary size.
\item Big integer values are clipped. 
More specifically, an integer in the range from -255 to 256 are preserved as it is, and other integers are clipped and substituted with the \texttt{\_\_HEX\_\_} symbol.
This also reduces the vocabulary size.
\end{itemize}

@@

\subsection{C Code Preprocessing}
% 232 words
The C source codes were parsed to AST by a Clang compiler \citep{clang} and its Python bindings together with pycparser \citep{pycparser}, 
because all of the tools are individually not sufficient for generating a dataset.
Then, the AST is converted to input data as follows:

\begin{itemize}
\item The variable-ary nodes and optional nodes were converted. 
For example, the \texttt{CompoundStatement} node has a variable number of statements. 
We converted such a list of nodes to the chains of the node \texttt{CONS} and the node \texttt{NIL}, as the list representation in a functional language.
Another example is the \texttt{if} statement, whose \texttt{else} clause is sometimes missing. 
We converted such optional nodes to the alternative of the \texttt{SOME} node and the \texttt{NONE} node, as the optional data structure in a functional language.
\item The variable names, function names, label names, or type names were renamed similarly to the addresses in the assembly code data.
\item There are four type of literals in the C language. 
An integer literal less than 256 was preserved, and an integer literal greater than 256 was clipped to 256.
The string, float and character literals were converted to the \texttt{\_\_LITERAL\_\_} symbol.

@@

\section{Evaluation Metrics}
We evaluated the quality of the results by two metrics, 
the edit distance ratio \citep{Motoneta} and the bilingual evaluation understudy (BLEU) \citep{BLEU}.

The edit distance ratio is a metric for for the correctness of the decompilation result.
Here, the edit distance ratio between two strings means the Levenshtein distance \citep{levensthein_dist} divided by the length of the longer string.
Ideally, the correctness of the decompilation result should be measured by the distance of the meaning of the program.
However the decompilation result is usually uncompilable, 
thus we used the edit distance of the source codes as an approximation of the correctness.
